Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-09 paper.tex
Rough out the Technology section.

\section{Technology}
\subsection{Rule-based story modules}
% excerpt of story model
% rete algorithm, cite Riedl, etc.
% programmable conflict resolution
\subsection{Simple dialog with AIML}
% eliza interactions, show fragment
\subsection{Web technologies}\label{sec:turtles}
% Javascript/HTML5
% offline caching
% turtles (almost) all the way down (turtlescript again)
% excerpt of TurtleScript

\section{Related work}\label{sec:related}
% Nell focused on Literacy
% olpc context here?
\cite{chang:tinkrbook}
%context: english literacy, far from teachers & schools
\subsection{The Sugar Learning Environment}
% Sugar
%% customization
%% path through the educational material?
\subsection{Narrative computing}
% Reidl
\subsection{Storytelling}
% TinkRBooks

>>>>

\section{Technology}
We have begun implementation of the Nell system.  In this section we
will describe the implementation of three of the most novel pieces
from the software engineering side.  These are not necessarily the
most novel-appearing from the user's perspective: for example, the
handwriting UI is still fairly unusual in consumer devices, but the
implementation of a separated-character handwriting recognition engine
using Hidden Markov Models is well-understood.
\textbf{SMOOTH THIS LANGUAGE}

\subsection{Rule-based story modules}
Narrative is the core of the Nell interface.  It is important that
Nell's story be modular and extensible.  We use a rule-based story
model with programmable conflict resolution to provide this ability.
Figure~\ref{fig:rules} shows an excerpt of the story module driving
the interaction described in the introduction to this paper.

XXX stories consist of elements.  all elements contain preconditions,
and actions.  cite PLANNER.

In principle, all rules for all elements of all stories are evaluated
continuously to determine the actions of the narrative agents.  This
is made computationally efficient using the Rete
algorithm~\cite{rete}.  The Rete algorithm requires that our
conditions are in Normal Form\footnote{BE PRECISE HERE}, but actions
can be arbitrarily complex.

New story modules can be incorporated by visiting an internet page in
connected deployments.  In a disconnected deployments, new stories
might be distributed once a year on USB sticks, or shared from your
friends.  In a classroom environment, a teacher can share stories for
the day's lesson at the start of class.

Story modules can be inherited and extended, as figure~\ref{fig:rules2}
shows.  This makes it easy to take an existing story and modify it to
better suit particular interests or a particular type of learner, or
just to add variety.  This can lead to conflicts: multiple versions of
a story, or multiple variant elements within a story, may have their
preconditions satisfied simultaneously.

We resolve conflicts...

We have not yet implemented an integrated story editor within Nell,
but we expect it to be roughly similar to the one implemented for
Wide Ruled, which showed good learnability by inexperienced
users~\cite{wideruled}.

% excerpt of story model
% rete algorithm, cite Riedl, etc.
% programmable conflict resolution
% cite the story editor from Wide Ruled

\subsection{Extensible dialog with AIML}
The child can interact with the Princess, and other agents within
Nell, directly.  Convincing discourse is a hard problem, but
reasonable approximations do not have to be difficult.  We leverage
the AIML language developed for AliceBot~\cite{xxx}, which has
won several Turing contests. \textbf{XXX more precise here XXX}
\footnote{Anecdotal evidence shows that interactions with Eliza-style
  responders can be rewarding: one woman spent hours each day
  conversing with the original Eliza.~\cite{RadioLab}}

% eliza interactions, show fragment
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.  Figure~\ref{fig:aiml}
illustrates how this can be used to implement a traditional
text adventure interaction for a particular story module.
The child can write instructions such as, ``go north'', ``go
east'', or ``open door'' for the agent.  These cause trigger events
which the activity responds to to move the agent through the
adventure---for example, by drawing the new location and causing
the agent to describe it.

% xx can the activity influence which commands are active?

\subsection{TurtleScript}\label{sec:turtles}
Nell's implementation leverages Web technologies, including HTML5,
WebGL, and JavaScript, with extensive use of offline caching.
Resources are named by URL, even when disconnected from the
internet.  This allows third parties to manage their own namespaces
when extending Nell, and simplifies updates to story modules and
the Nell system itself.
% if the kid's going to learn some software stack, this seem like a
% fine one to learn

As far as possible, Nell is implemented in JavaScript.  We have
developed a block-oriented view of JavaScript, which we call
TurtleScript~\cite{turtlescript}.  Combined with subsets of JavaScript
for beginners, this allows us to use a single implementation language
which scales from logo-like tasks all the way through the
implementation of the entire Nell system itself.  TurtleScript's block
view of JavaScript is isomorphic to the standard textual
representation of code.  We have begun implementation of a full
programming environment for TurtleScript which is naturally-suited to
tablet machines.

% XXX refer to TurtleScript excerpt in figure somewhere here.

Our goal is to allow the child to eventually dig as far into Nell's
implementation as their capabilities allow.  Although the operating system and
the browser itself are inaccessible from the web platform, we can
provide full access above that level.  The existence of meta-circular
interpreters for JavaScript~\cite{narcissus} implies that we can
expose the system all the way down to the programming language
implementation.  Via TurtleScript, this code is all visible in a
friendly block-based form.

% Javascript/HTML5
% offline caching
% turtles (almost) all the way down (turtlescript again)
% excerpt of TurtleScript

\section{Related work}\label{sec:related}

% Nell focused on Literacy
% olpc context here?
\cite{chang:tinkrbook}
%context: english literacy, far from teachers & schools
\subsection{The Sugar Learning Environment}
% Sugar
%% customization
%% path through the educational material?
\subsection{Narrative computing}
% Reidl
\subsection{Storytelling}
% TinkRBooks
